---
title: "P(XYZ) Table"
author: "Rose Wang"
date: "7/20/2019"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
setwd("/Users/rosewang/Downloads/RCompSci")
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(jsonlite)
library(httr)
library(scatterplot3d)
library(lattice)
library(mefa)
library(glmnet)
library(knitr)
```

## Univariate Counts

In order to verify the P(XYZ) table, we will first get the Univariate Counts of three variables (Total ED/Inpatient Visits, Average Daily PM 2.5 Exposure, Major Roadway Highway Exposure) from the ICEES API. We will use the parameters of: Cohort 22 (all patients), year 2010, version 1.0.0 of ICEES data, and patient level data. We will also be using the Feature-rich cohort Discovery functionality.

```{r univariate}
url <- "https://icees.renci.org/1.0.0/patient/2010/cohort/COHORT%3A22/features"
uni <- GET(url,
          accept_json(),
          content_type("application/json"),
          encode = "json")
uni.text <- content(uni, "text")
uni.json <- fromJSON(uni.text, flatten = TRUE)
uni.json[["terms and conditions"]] <- NULL
uni.json[["version"]] <- NULL
```

We will now convert the data from JSON form to a data frame by parsing the values into one big data frame. We will specifically be looking into Total ED/Inpatient Visits, Average Daily PM 2.5 Exposure, Major Roadway Highway Exposure.

```{r unidataframe}
featureNumber <- c(17, 29, 30)
featureNames <- uni.json[["return value"]][["feature.feature_name"]]
featureNames <- as.vector(c(featureNames[17], featureNames[29], featureNames[30]))
for(i in 1:3){
  assign(paste("freq",uni.json[["return value"]][["feature.feature_name"]][[featureNumber[i]]],sep = ""), data.frame(c(uni.json[["return value"]][["feature.feature_qualifiers"]][[featureNumber[i]]][["value"]], "Total"), frequency = c(uni.json[["return value"]][["feature_matrix"]][[featureNumber[i]]][["frequency"]], sum(uni.json[["return value"]][["feature_matrix"]][[featureNumber[i]]][["frequency"]]))))
  if(i==2){
    colnames(freqMajorRoadwayHighwayExposure)[1] <- paste("",uni.json[["return value"]][["feature.feature_name"]][[featureNumber[i]]],sep = "")
  } else if(i==1){
    colnames(freqMaxDailyPM2.5Exposure_qcut)[1] <- paste("",uni.json[["return value"]][["feature.feature_name"]][[featureNumber[i]]],sep = "")
  } else {
    colnames(freqTotalEDInpatientVisits)[1] <- paste("",uni.json[["return value"]][["feature.feature_name"]][[featureNumber[i]]],sep = "")
  }
}
```

Finished Tables:

``` {r unidisplay, echo=FALSE}
kable(freqMajorRoadwayHighwayExposure, caption = "Major Roadway Highway Exposure Univariate Table")
kable(freqMaxDailyPM2.5Exposure_qcut, caption = "Max Daily PM2.5 Exposure Univariate Table")
kable(freqTotalEDInpatientVisits, caption = "Total ED/Inpatient Visits Univariate Table")
```

## Bivariate Counts

In order to verify the P(XYZ) table, we also have to get the Bivariate Counts of the three bivariate combinations of the three variables (Total ED/Inpatient Visits, Average Daily PM 2.5 Exposure, Major Roadway Highway Exposure) from the ICEES API. We will use the parameters of: Cohort 22 (all patients), year 2010, version 1.0.0 of ICEES data, and patient level data. We will also be using the Hypothesis Driven NxN Feature Association Functionality.

We will start off with the Total ED/Inpatient Visits vs. Max Daily PM2.5 Exposures Table. 

```{r EDVisits-PM2.5}
PM2.5data <- POST("https://icees.renci.org/1.0.0/patient/2010/cohort/COHORT%3A22/feature_association2",
                     body ='{
                     "feature_a": {
                     "TotalEDInpatientVisits": [
                     {
                     "operator": "=",
                     "value": 0
                     }, {
                     "operator": "=",
                     "value": 1
                     }, {
                     "operator": "=",
                     "value": 2
                     },{
                     "operator": "=",
                     "value": 3
                      },{
                     "operator": "=",
                    "value": 4
                     }, {
                     "operator": "=",
                     "value": 5
                     }, {
                     "operator": "=",
                     "value": 6
                     }, {
                     "operator": "=",
                     "value": 7
                     }, {
                     "operator": "=",
                     "value": 8
                     }, {
                     "operator": "=",
                      "value": 9
                     }
                     ]
                     },
                     "feature_b": {
                     "MaxDailyPM2.5Exposure_qcut": [
                     {
                     "operator": "=",
                     "value": 1
                     },
                     {
                     "operator": "=",
                     "value": 2
                     },
                     {
                     "operator": "=",
                     "value": 3
                     },
                     {
                     "operator": "=",
                    "value": 4
                     },
                     {
                     "operator": "=",
                     "value": 5
                     }
                     ]
                     }
                     }',
          accept_json(),
          content_type("application/json"),
          encode = "json")
PM2.5datatext <- content(PM2.5data, "text")
#
PM2.5data_json <- fromJSON(PM2.5datatext, flatten = TRUE)
PM2.5data_json$"terms and conditions" <- NULL
PM2.5data_json$"version" <- NULL
features.df <- data.frame("EDvisits" = c(0,1,2,3,4,5,6,7,8,9))
PM2.5.df <- data.frame("frequency" = c(), "column_percentage" = c(), "row_percentage" = c(), "total_percentage" = c(),"EDvisits" = c(), "MaxDailyPM2.5Exposure_qcut" = c())
for(i in 1:5){
  PM2.5feature_matrix.df <- as.data.frame(PM2.5data_json$"return value"$feature_matrix[[i]])
  t <- PM2.5feature_matrix.df %>%
    bind_cols(features.df, PM2.5feature_matrix.df) %>%
    subset(select = -c(frequency1, column_percentage1, row_percentage1, total_percentage1)) %>%
    mutate(MaxDailyPM2.5Exposure_qcut = i)
  t <- as.data.frame(t)
  PM2.5.df <- rbind(PM2.5.df, t)
}
PM2.5.df <- PM2.5.df[c(5,6,1,2,3)]
```

Next, we will pull the Total ED/Inpatient Visits vs. Major Roadway Highway Exposures data. 

``` {r EDVisits-MajRoadway}
MajRoadwaydata <- POST("https://icees.renci.org/1.0.0/patient/2010/cohort/COHORT%3A22/feature_association2",
                  body ='{
                  "feature_a": {
                  "TotalEDInpatientVisits": [
                     {
                  "operator": "=",
                  "value": 0
                  }, {
                  "operator": "=",
                  "value": 1
                  }, {
                  "operator": "=",
                  "value": 2
                  },{
                  "operator": "=",
                  "value": 3
                  },{
                  "operator": "=",
                  "value": 4
                  }, {
                  "operator": "=",
                  "value": 5
                  }, {
                  "operator": "=",
                  "value": 6
                  },
                  {
                  "operator": "=",
                  "value": 7
                  }, {
                  "operator": "=",
                  "value": 8
                  }, {
                  "operator": "=",
                  "value": 9
                  }
                  ]
                  },
                  "feature_b": {
                  "MajorRoadwayHighwayExposure": [
                  {
                  "operator": "=",
                  "value": 1
                  },
                  {
                  "operator": "=",
                  "value": 2
                  },
                  {
                  "operator": "=",
                  "value": 3
                  },
                  {
                  "operator": "=",
                  "value": 4
                  },
                  {
                  "operator": "=",
                  "value": 5
                  },
                  {
                  "operator": "=",
                  "value": 6
                  }
                  ]
                  }
                  }',
          accept_json(),
          content_type("application/json"),
          encode = "json")
MajRoadwaydatatext <- content(MajRoadwaydata, "text")
#
MajRoadwaydata_json <- fromJSON(MajRoadwaydatatext, flatten = TRUE)
MajRoadwaydata_json$"terms and conditions" <- NULL
MajRoadwaydata_json$"version" <- NULL
features.df <- data.frame("EDvisits" = c(0,1,2,3,4,5,6,7,8,9))
MajRoadway.df <- data.frame("frequency" = c(), "column_percentage" = c(), "row_percentage" = c(), "total_percentage" = c(),"EDvisits" = c(), "MajorRoadwayHighwayExposure" = c())
for(i in 1:6){
  MajRoadwayfeature_matrix.df <- as.data.frame(MajRoadwaydata_json$"return value"$feature_matrix[[i]])
  t <- MajRoadwayfeature_matrix.df %>%
    bind_cols(features.df, MajRoadwayfeature_matrix.df) %>%
    subset(select = -c(frequency1, column_percentage1, row_percentage1, total_percentage1)) %>%
    mutate(MajorRoadwayHighwayExposure = i)
  t <- as.data.frame(t)
  MajRoadway.df <- rbind(MajRoadway.df, t)
}
MajRoadway.df <- MajRoadway.df[c(5,6,1,2,3)]
```

Finally, we will pull the Major Roadway Highway Exposures vs. Max DailyPM2.5 Exposures table. 

```{r MajRoadway-PM2.5}
MajPM2.5data_ <- POST("https://icees.renci.org/1.0.0/patient/2010/cohort/COHORT%3A22/feature_association2",
                      body ='{
                      "feature_a": {
                      "MajorRoadwayHighwayExposure": [
                      {
                      "operator": "=",
                      "value": 1
                      },
                      {
                      "operator": "=",
                      "value": 2
                      },
                      {
                      "operator": "=",
                      "value": 3
                      },
                      {
                      "operator": "=",
                      "value": 4
                      },
                      {
                      "operator": "=",
                      "value": 5
                      },
                      {
                      "operator": "=",
                      "value": 6
                      }
                      ]
                      },
                      "feature_b": {
                      "MaxDailyPM2.5Exposure_qcut": [
                      {
                      "operator": "=",
                      "value": 1
                      },
                      {
                      "operator": "=",
                      "value": 2
                      },
                      {
                      "operator": "=",
                      "value": 3
                      },
                      {
                      "operator": "=",
                      "value": 4
                      },
                      {
                      "operator": "=",
                      "value": 5
                      }
                      ]
                      }
                      }',
          accept_json(),
          content_type("application/json"),
          encode = "json")
MajPM2.5data_text <- content(MajPM2.5data_, "text")
#
MajPM2.5data_json <- fromJSON(MajPM2.5data_text, flatten = TRUE)
MajPM2.5data_json$"terms and conditions" <- NULL
MajPM2.5data_json$"version" <- NULL
features.df <- data.frame("MajorRoadwayHighwayExposure" = c(1,2,3,4,5,6))
MajPM2.5.df <- data.frame("frequency" = c(), "column_percentage" = c(), "row_percentage" = c(), "total_percentage" = c(),"MajorRoadwayHighwayExposure" = c(), "MaxDailyPM2.5Exposure_qcut" = c())
for(i in 1:5){
  MajPM2.5feature_matrix.df <- as.data.frame(MajPM2.5data_json$"return value"$feature_matrix[[i]])
  t <- MajPM2.5feature_matrix.df %>%
    bind_cols(features.df, MajPM2.5feature_matrix.df) %>%
    subset(select = -c(frequency1, column_percentage1, row_percentage1, total_percentage1)) %>%
    mutate(MaxDailyPM2.5Exposure_qcut = i)
  t <- as.data.frame(t)
  MajPM2.5.df <- rbind(MajPM2.5.df, t)
}
MajPM2.5.df <- MajPM2.5.df[c(5,6,1,2,3)]
```

Finished Tables:

``` {r bidisplay, echo=FALSE}
kable(MajRoadway.df, caption = "Total ED/Inpatient Visits vs. Major Roadway Highway Exposure Bivariate Table")
kable(PM2.5.df, caption = "Total ED/Inpatient Visits vs. Max Daily PM2.5 Exposure Bivariate Table")
kable(MajPM2.5.df, caption = "Major Roadway Highway Exposure vs. Max Daily PM2.5 Exposure Bivariate Table")
```

## Trivariate Counts

Finally, we will create the P(XYZ) table. In order to do so, we will be narrowing the cohort of each of our calls to the API to one bin of the Total ED/Inpatient Visits feature. Thus, we first have to create each of the cohorts that we will be using. 

``` {r cohortcreation}
cohort <- c()
urls <- "https://icees.renci.org/1.0.0/patient/2010/cohort"
for(i in 0:9){
  bodys <- paste('{"TotalEDInpatientVisits":{"operator":"=","value":',i,'}}', sep = "")
  data <- POST(urls, body = as.character(bodys), accept_json(), content_type("application/json"), encode = "json")
  data.text <- content(data, "text")
  data.json <- fromJSON(data.text, flatten = TRUE)
  c <- data.json[["return value"]][["cohort_id"]]
  cohort <- c(cohort, c)
}
cohort <- strsplit(cohort, "COHORT:")
```

Next, we can use the cohorts we just created to pull the bivariate distribution between Major Roadway Highway Exposure and Max Daily PM2.5 Exposure. We will use the parameters of: varying cohorts, year 2010, version 1.0.0 of ICEES data, and patient level data. We will also be using the Hypothesis Driven NxN Feature Association Functionality.

``` {r P(XYZ)table}
total <- data.frame("TotalEDInpatientVisits" = c(), "MajorRoadwayHighwayExposure" = c(), "MaxDailyPM2.5Exposure_qcut" = c(), "frequency" = c(), "column_percentage" = c(), "row_percentage" = c(), "total_percentage" = c())
for(i in 1:10){
  urls <- paste('https://icees.renci.org/1.0.0/patient/2010/cohort/COHORT%3A',cohort[[i]][2],'/feature_association2', sep = "")
  Bidata <- POST(as.character(urls),
                        body ='{
                        "feature_a": {
                        "MajorRoadwayHighwayExposure": [{
                        "operator": "=",
                        "value": 1
                        },{
                        "operator": "=",
                        "value": 2
                        },{
                        "operator": "=",
                        "value": 3
                        },{
                        "operator": "=",
                        "value": 4
                        },{
                        "operator": "=",
                        "value": 5
                        },
                        {
                        "operator": "=",
                        "value": 6
                        }]},
                        "feature_b": {
                        "MaxDailyPM2.5Exposure_qcut": [{
                        "operator": "=",
                        "value": 1
                        },{
                        "operator": "=",
                        "value": 2
                        },{
                        "operator": "=",
                        "value": 3
                        },{
                        "operator": "=",
                        "value": 4
                        },{
                        "operator": "=",
                        "value": 5
                        }]}}',
          accept_json(),
          content_type("application/json"),
          encode = "json")
  Bidata_text <- content(Bidata, "text")
  
  # convert into data frame
  Bidata_json <- fromJSON(Bidata_text, flatten = TRUE)
  Bidata_json[["terms and conditions"]] <- NULL
  Bidata_json[["version"]] <- NULL
  Bi.df <- data.frame("frequency" = c(), "column_percentage" = c(), "row_percentage" = c(), "total_percentage" = c(),"TotalEDInpatientVisits" = c(), "MajorRoadwayHighwayExposure" = c(), "MaxDailyPM2.5Exposure_qcut" = c())
  for(n in 1:5){
    Bifeature_matrix.df <- as.data.frame(Bidata_json[["return value"]][["feature_matrix"]][[n]])
    t <- Bifeature_matrix.df %>%
      mutate(TotalEDInpatientVisits = i-1, MajorRoadwayHighwayExposure = 1:6, MaxDailyPM2.5Exposure_qcut = n )
    Bi.df <- rbind(Bi.df, t)
  }
  total <- rbind(total, Bi.df)
}
total <- total[,c("TotalEDInpatientVisits", "MajorRoadwayHighwayExposure", "MaxDailyPM2.5Exposure_qcut", "frequency")]
```

Finished Trivariate Table:

```{r tridisplay, echo=FALSE}
kable(total, caption = "Total ED/Inpatient Visits vs. Major Roadway Highway Exposure vs. Max Daily PM2.5 Exposure Trivariate Table")
```

For a final plot, we also ended up simulating the Major roadway Highway Exposure and Max Daily PM2.5 Exposure data using a random uniform distribution. We then plotted and made a regression for the trivariate simulated data.

```{r simulateplotregress, echo=FALSE}
PM2.5bounds <- data.frame("PM2.5bounds" = c(6.768784523, 42.0219841, 46.209636688, 47.062618256, 51.719459534, 114.94243622))
Roadwaybounds <- data.frame("Roadwaybounds" = c(0,50,100,200,300,500))
simulated.df <- data.frame("TotalEDInpatientVisits" = c(), "MajorRoadwayHighwayExposure" = c(), "MaxDailyPM2.5Exposure_qcut" = c(), "sMajorRoadwayHighwayExposure" = c(), "sMaxDailyPM2.5Exposure" = c())
for (i in 1:5){
  simPM2.5 <- total %>%
    filter(MaxDailyPM2.5Exposure_qcut == i)
  for (n in 1:5){
    simRoadway <- simPM2.5 %>%
      filter(MajorRoadwayHighwayExposure == n)
    freq <- simRoadway$frequency
    t <- c()
    a <- c()
    c <- c()
    for (l in 1:10){
      sMajorRoadwayHighwayExposure <- runif(as.numeric(freq[l]), min=as.numeric(Roadwaybounds$Roadwaybounds[n]), max=as.numeric(Roadwaybounds$Roadwaybounds[n+1]))
      sMajorRoadwayHighwayExposure <- as.data.frame(sMajorRoadwayHighwayExposure)
      t <- rbind(t, sMajorRoadwayHighwayExposure)
      sMaxDailyPM2.5Exposure <- runif(as.numeric(freq[l]), min=as.numeric(PM2.5bounds$PM2.5bounds[i]), max=as.numeric(PM2.5bounds$PM2.5bounds[i+1]))
      sMaxDailyPM2.5Exposure <- as.data.frame(sMaxDailyPM2.5Exposure)
      a <- rbind(a, sMaxDailyPM2.5Exposure)
      features <- data.frame("TotalEDInpatientVisits" = c(l-1),"MajorRoadwayHighwayExposure" = c(as.numeric(n)), "MaxDailyPM2.5Exposure_qcut" = c(as.numeric(i)))
      features <- as.data.frame(features)
      features <- rep(features, times = freq[l])
      features <- as.data.frame(features)
      c <- rbind(c,features)
    }
    filler <- bind_cols(c, t, a)
    simulated.df <- rbind(simulated.df, filler)
  }
}
simulated.df <- simulated.df[,c("TotalEDInpatientVisits","sMajorRoadwayHighwayExposure","sMaxDailyPM2.5Exposure")]

# Scatter Plot and Regression
s3d <- scatterplot3d(x = simulated.df$sMajorRoadwayHighwayExposure,
                     y = simulated.df$sMaxDailyPM2.5Exposure,
                     z = simulated.df$TotalEDInpatientVisits,
                     pch = 20,
                     cex.symbols = .5,
                     color = "#E69F00",
                     xlab = "Major Roadway Highway Exposure",
                     ylab = "Maximum Daily PM2.5 Exposure",
                     zlab = "Number of Inpatient Visits",
                     angle = 20,
                     box = FALSE)
regression <- lm(simulated.df$TotalEDInpatientVisits~
                  simulated.df$sMajorRoadwayHighwayExposure+
                  simulated.df$sMaxDailyPM2.5Exposure)
s3d$plane3d(regression)
```

## Comparisons

Since ICEES doesn't output data if the value is less than 10, some data will be lost when decreasing the size of different cohorts. Thus, it is necessary to see how much data was lost. We will see what and how much data is lost by creating comparison tables for each bin level of each feature. In order to do so, we will use the the filter function to find the total frequency of each bin level for each feature in the univariate, bivariate, and trivariate features. 

First, we will focus on Total ED/Inpatient Visits. 

``` {r totalEDVisits}
totalEDVisits <- data.frame(TotalEDInpatientVisits = c(0,1,2,3,4,5,6,7,8,9), UnivariateFreq = c(freqTotalEDInpatientVisits$frequency[1:10]))
MajRoadwayFreq <- data.frame(MajRoadwayFreq = c())
PM2.5Freq <- data.frame(PM2.5Freq = c())
TriFreq <- data.frame(TriFreq = c())
for(i in 0:9){
  a <- MajRoadway.df %>%
    filter(EDvisits == i)
  b <- PM2.5.df %>%
    filter(EDvisits == i)
  c <- total %>%
    filter(TotalEDInpatientVisits == i)
  t <- data.frame(MajRoadwayFreq = c(sum(a$frequency)))
  s <- data.frame(PM2.5Freq = c(sum(b$frequency)))
  r <- data.frame(TriFreq = c(sum(c$frequency)))
  MajRoadwayFreq <- rbind(MajRoadwayFreq, t)
  PM2.5Freq <- rbind(PM2.5Freq, s)
  TriFreq <- rbind(TriFreq, r)
}
totalEDVisits <- cbind(totalEDVisits, MajRoadwayFreq, PM2.5Freq, TriFreq)
```

Next, we will look into Major Roadway Highway Exposures. 

``` {r totalMajRoadway}
totalMajRoadway <- data.frame(MajorRoadwayHighwayExposure = c(1,2,3,4,5,6), UnivariateFreq = c(freqMajorRoadwayHighwayExposure$frequency[1:6]))
EDVisitsFreq <- data.frame(EDVisitsFreq = c())
PM2.5Freq <- data.frame(PM2.5Freq = c())
TriFreq <- data.frame(TriFreq = c())
for(i in 1:6){
  a <- MajRoadway.df %>%
    filter(MajorRoadwayHighwayExposure == i)
  b <- MajPM2.5.df %>%
    filter(MajorRoadwayHighwayExposure == i)
  c <- total %>%
    filter(MajorRoadwayHighwayExposure == i)
  t <- data.frame(EDVisitsFreq = c(sum(a$frequency)))
  s <- data.frame(PM2.5Freq = c(sum(b$frequency)))
  r <- data.frame(TriFreq = c(sum(c$frequency)))
  EDVisitsFreq <- rbind(EDVisitsFreq, t)
  PM2.5Freq <- rbind(PM2.5Freq, s)
  TriFreq <- rbind(TriFreq, r)
}
totalMajRoadway <- cbind(totalMajRoadway, EDVisitsFreq, PM2.5Freq, TriFreq)
```

Finally, we will compute Max Daily PM2.5 Exposures. 

``` {r totalPM2.5}
totalPM2.5 <- data.frame(MaxDailyPM2.5Exposure_qcut = c(1,2,3,4,5), UnivariateFreq = c(freqMaxDailyPM2.5Exposure_qcut$frequency[1:5]))
EDVisitsFreq <- data.frame(EDVisitsFreq = c())
MajRoadwayFreq <- data.frame(MajRoadwayFreq = c())
TriFreq <- data.frame(TriFreq = c())
for(i in 1:5){
  a <- PM2.5.df %>%
    filter(MaxDailyPM2.5Exposure_qcut == i)
  b <- MajPM2.5.df %>%
    filter(MaxDailyPM2.5Exposure_qcut == i)
  c <- total %>%
    filter(MaxDailyPM2.5Exposure_qcut == i)
  t <- data.frame(EDVisitsFreq = c(sum(a$frequency)))
  s <- data.frame(MajRoadwayFreq = c(sum(b$frequency)))
  r <- data.frame(TriFreq = c(sum(c$frequency)))
  EDVisitsFreq <- rbind(EDVisitsFreq, t)
  MajRoadwayFreq <- rbind(MajRoadwayFreq, s)
  TriFreq <- rbind(TriFreq, r)
}
totalPM2.5 <- cbind(totalPM2.5, EDVisitsFreq, MajRoadwayFreq, TriFreq)

```

Finished Count Comparison Tables:

``` {r comparisondisplay, echo=FALSE}
kable(totalEDVisits, caption = "Univariate, Bivariate, and Trivariate Count Comparison - Total ED/Inpatient Visits")
kable(totalMajRoadway, caption = "Univariate, Bivariate, and Trivariate Count Comparison - Major Roadway Highway Exposures")
kable(totalPM2.5, caption = "Univariate, Bivariate, and Trivariate Count Comparison - Max Daily PM2.5 Exposures")
```