---
title: "Make Univariate Tables"
author: "Perry Haaland"
date: "03/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(jsonlite)
library(httr)
```

## Introduction

Eventually we are going to extract a three-way table and the three corresponding two-way tables using the OpenAPI, and the univariate counts for each variable separately. Because we are using the OpenAPI, we are extracting pre-binned data. The extracted values are also subject the privacy constraints of the system in terms of reporting low frequency cell counts.

The univariate counts for each variable are the univariate marginal distribution across the bins. The two-way tables are the bivariate marginal distributions. The three-way table, which is the full multivariate distribution is constrained by its marginal distributions, so we can use those to check on the correctness of the OpenAPI results.


We use the `json` format of the interface. The data set that we aer extracting from is

* Cohort 22 (all patients),
* 2010 data,
* Version 1.0.0 of ICEES.

In this file, we extract the univariate tables for each feature. The output of this file is the data frame `uni_all_df`.

## Extracting the univariate counts using the OpenAPI

We first get the univariate counts for all features (variables) the ICEES database using the ICEES API. Note that we first specify the ICEES URL and then we use the `GET` functionality from the R package `httr`. Then we use functions from the package `jsonlite` to convert from JSON to R.


```{r univariate, message=FALSE}
## no need to see uninformative messages in the markdown document
url <- "https://icees.renci.org/1.0.0/patient/2010/cohort/COHORT%3A22/features"
## extract the data in json format
univariate.call <- httr::GET(url,
          accept_json(),
          content_type("application/json"),
          encode = "json")
## the next line of code extracts just the content as a character string
univariate.text <- httr::content(univariate.call, "text")
## convert the output to JSON format
## note that this contains all of the variables in the dataset
univariate.list <- jsonlite::fromJSON(univariate.text, flatten = TRUE)$`return value`
```

Note that the current data object, `univariate.list`, has the univariate binned distribution for all of the features.

```{r}
names(univariate.list)
```

The first component of the output are the frequencies and percentages in each bin.

```{r}
univariate.list$feature_matrix[[1]]
```

The second component of the output contains the names of the features. We show the first five names here.

```{r}
cat("Number of Features = ", length(univariate.list$feature.feature_name),"\n")
univariate.list$feature.feature_name[1:5]
```

The third component provides the bin identifications. This may be a range or a single value. The bin identifiers for the first variable `AgeStudyStart` are shown below. This feature has been reported in 6 bins, based on age ranges.

```{r, message=FALSE}
univariate.list$feature.feature_qualifiers[[1]]
```

The bins may also correspond to single values as for the second feature, `Sex`.


```{r, message=FALSE}
univariate.list$feature.feature_qualifiers[[2]]
```


We will now extract the data from JSON form to a data frame by parsing the values into one big data frame. 


```{r}
## the following function combines the results for one feature
make_df <- function(i) {
  data.frame(feature=univariate.list$feature.feature_name[i],
             univariate.list$feature.feature_qualifiers[i],
              univariate.list$feature_matrix[[i]])
}
## this code loops through the entire list and returns a single data frame
## pdh: should remember how to do this more elegantly!
for(i in 1:length(univariate.list[[1]])) {
  if(i == 1) {
    uni_all_df <- make_df(i)
  } else {
    uni_all_df <- rbind(uni_all_df,make_df(i))
  }
}
```

This data frame is called `uni_all_df`. Here are the results for the first two features.

```{r}
uni_all_df[1:8,]
```

The size of the data frame is 

```{r}
dim(uni_all_df)
```

The number of features is

```{r}
length(unique(uni_all_df$feature))
```

## Make a smaller table with just the three relevant variables

We select the three features of interest and rename the variables with simpler names for use later.

* `EDVisits` = Total ED/Inpatient Visits, 
* `MaxDailyExposure` = Maximum Daily PM 2.5 Exposure (qcut), 
* `RoadwayExposure` = Major Roadway Highway Exposure.

```{r}
uni_df <- uni_all_df %>% 
  filter(str_detect(feature, 'MaxDailyPM2.5Exposure_qcut|MajorRoadwayHighwayExposure|TotalEDInpatientVisits')) %>%
  select(Feature = feature, Value = value, Frequency = frequency) %>%
  mutate(Feature = recode(Feature, TotalEDInpatientVisits = "EDVisits",
                          MajorRoadwayHighwayExposure = "RoadwayExposure",
                          MaxDailyPM2.5Exposure_qcut = "MaxDailyExposure"))
uni_df
  
```



## Save the resulting data frames

```{r}
save(uni_all_df, file = "RData/uni_all_df.RData")
save(uni_df, file = "RData/uni_df.RData")
```


```

