---
title: "Make Trivariate Table"
author: "Perry Haaland"
date: "03/05/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(jsonlite)
library(httr)
```

## Trivariate Counts

In this file we make three-way, or trivariate table. In previous files we created the univariate table and the three bivariate tables for Roadway Exposure, Emergency Room Visits, and PM2.5 exposure. In order create the three-way table, we narrow the cohort of each of our calls to the API to one bin of the Total ED/Inpatient Visits feature. Thus, we first have to create each of the cohorts that we will be using. 

```{r cohortcreation, message=FALSE}
tri_cohorts <- c()
urls <- "https://icees.renci.org/1.0.0/patient/2010/cohort"
for(i in 0:9){
  tri_body_tmp <- paste('{"TotalEDInpatientVisits":{"operator":"=","value":', i,'}}', sep = "")
  tri_call_tmp <- httr::POST(urls, 
                             body = as.character(tri_body_tmp), 
                     accept_json(), 
                     content_type("application/json"), 
                     encode = "json")
  tri_text_tmp <- httr::content(tri_call_tmp, "text")
  tri_cohort <- jsonlite::fromJSON(tri_text_tmp, flatten = TRUE)$`return value`$cohort_id
  tri_cohort = strsplit(tri_cohort, "COHORT:")[[1]][2]
  tri_cohorts <- c(tri_cohorts, tri_cohort)
}
```

Next, we can use the cohorts we just created to pull the bivariate distribution between Major Roadway Highway Exposure and Max Daily PM2.5 Exposure. We will use the parameters of: varying cohorts, year 2010, version 1.0.0 of ICEES data, and patient level data. We will also be using the Hypothesis Driven NxN Feature Association Functionality.

```{r trivariate table, message=FALSE}
for(i in 1:10){
  cohort_url <- paste('https://icees.renci.org/1.0.0/patient/2010/cohort/COHORT%3A',
                tri_cohorts[i],
                '/feature_association2', 
                sep = "")
  cohort_call <- httr::POST(as.character(cohort_url),
                        body ='{
                        "feature_a": {
                        "MajorRoadwayHighwayExposure": [{
                        "operator": "=",
                        "value": 1
                        },{
                        "operator": "=",
                        "value": 2
                        },{
                        "operator": "=",
                        "value": 3
                        },{
                        "operator": "=",
                        "value": 4
                        },{
                        "operator": "=",
                        "value": 5
                        },
                        {
                        "operator": "=",
                        "value": 6
                        }]},
                        "feature_b": {
                        "MaxDailyPM2.5Exposure_qcut": [{
                        "operator": "=",
                        "value": 1
                        },{
                        "operator": "=",
                        "value": 2
                        },{
                        "operator": "=",
                        "value": 3
                        },{
                        "operator": "=",
                        "value": 4
                        },{
                        "operator": "=",
                        "value": 5
                        }]}}',
          accept_json(),
          content_type("application/json"),
          encode = "json")
  cohort_text <- httr::content(cohort_call, "text")
  cohort_list <- jsonlite::fromJSON(cohort_text, flatten = TRUE)$`return value`
  ## have to hard code these values
  EDVisits = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9)
  MaxRoadway = c(1, 2, 3, 4, 5, 6)
  MaxExposure = c(1, 2, 3, 4, 5)
  for(j in 1:5){
    temp_df <- data.frame(EDVisits = EDVisits[i],
                          RoadwayExposure = MaxRoadway, 
                          MaxExposure = MaxExposure[j],
                          Frequency = cohort_list$feature_matrix[[j]]$frequency)
    if(j == 1) {
      cohort_df = temp_df
    } else {
      cohort_df <- rbind(cohort_df,temp_df)
    }
  }
  if(i == 1){
    tri_df = cohort_df
  } else {
    tri_df = rbind(tri_df,cohort_df)
  }
  }
  dim(tri_df)
```

The table looks as follows:

```{r}
tri_df
```

The total number of subject reported in the trivariate table is as follows:

```{r}
sum(tri_df$Frequency)
```

## Save the resulting data frame


```{r}
save(tri_df, file = "RData/tri_df.RData")
```


