---
title: "Validation of Tables"
author: "Haaland"
date: "03/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```


## Load the tables

In files 2-4, we created and saved the univariate, bivariate and trivariate tables for the variables:

* `EDVisits` = Total ED/Inpatient Visits, 
* `MaxPMExposure` = Maximum Daily PM 2.5 Exposure (qcut), 
* `RoadwayExposure` = Major Roadway Highway Exposure.

The three pairwise combinations of binary tables are labeled as follows:

* `bivar_EDvsME_df`: `EDVisits` versus `MaxPMExposure`
* `bivar_EDvsRE_df`: `EDVisits` versus `RoadwayExposure`
* `bivar_REvsME_df`: `RoadwayExposure` versus `MaxPMExposure`

We start by loading the data.

```{r}
load("RData/uni_df.RData")
load("RData/bivar_EDvsME_df.RData")
load("RData/bivar_EDvsRE_df.RData")
load("RData/bivar_REvsME_df.RData")
load("RData/tri_df.RData")
```



## Brief examination of bivariate tables

## Compare bivariate marginals to univariate

To verify the correctness of the data, we can compare the marginal distributions from the bivariate tables to the univariate distributions.

### EDVisits

First we consider the univariate and marginal distributions for EDVisits. We calculate the marginal distributions from the bivariate and trivariate tables and compare to the univariate values. The results are be the same.

```{r}
bv1_df <- bivar_EDvsME_df %>%
  group_by(EDVisits) %>%
  summarize(bv1Frequency = sum(Frequency)) %>%
  select(vb1Value = EDVisits, bv1Frequency)
bv2_df <- bivar_EDvsRE_df %>%
  group_by(EDVisits) %>%
  summarize(bv2Frequency = sum(Frequency)) %>%
  select(vb2Value = EDVisits, bv2Frequency)
tr1_df <- tri_df %>%
  group_by(EDVisits) %>%
  summarize(triFrequency = sum(Frequency)) %>%
  select(truValue = EDVisits, triFrequency)
un_df <- uni_df %>%
  filter(Feature == "EDVisits",
         as.numeric(Value) < 10) %>%
  rename(uniFrequency = Frequency)
comp1_df <- cbind(un_df,bv1_df,bv2_df,tr1_df) %>%
  select(Feature, Value, uniFrequency, bv1Frequency, bv2Frequency, triFrequency)
comp1_df
```

We can see by inspection, that the results look correct. We test that formally below:

```{r}
comp1_df %>%
  transmute(Feature = Feature, Value = Value, agree = ((uniFrequency == bv1Frequency) & 
                                                         (uniFrequency == bv2Frequency) &
                                                        (uniFrequency == triFrequency)))
```

### RoadwayExposure

First we consider the univariate and marginal distributions for RoadwayExposure. We calculate the marginal distributions from the bivariate and trivariate tables and compare to the univariate values. In this comparison, we see differences among the data extraction approaches. I suggest that it matters which variable is specified as "a" in the call to the OpenAPI.

```{r}
bv1_df <- bivar_EDvsRE_df %>%
  group_by(RoadwayExposure) %>%
  summarize(bv1Frequency = sum(Frequency)) %>%
  select(vb1Value = RoadwayExposure, bv1Frequency)
bv2_df <- bivar_REvsME_df %>%
  group_by(RoadwayExposure) %>%
  summarize(bv2Frequency = sum(Frequency)) %>%
  select(vb2Value = RoadwayExposure, bv2Frequency)
tr1_df <- tri_df %>%
  group_by(RoadwayExposure) %>%
  summarize(triFrequency = sum(Frequency)) %>%
  select(truValue = RoadwayExposure, triFrequency)
un_df <- uni_df %>%
  filter(Feature == "RoadwayExposure") %>%
  rename(uniFrequency = Frequency)
comp2_df <- cbind(un_df,bv1_df,bv2_df,tr1_df) %>%
  select(Feature, Value, uniFrequency, bv1Frequency, bv2Frequency, triFrequency)
comp2_df
```

### MaxPMExposure

First we consider the univariate and marginal distributions for MaxPMExposure. We calculate the marginal distributions from the bivariate and trivariate tables and compare to the univariate values. In this comparison, we see differences among the data extraction approaches. Interestingly, the univariate frequency agrees with the REvsME table.

```{r}
bv1_df <- bivar_EDvsME_df %>%
  group_by(MaxPMExposure) %>%
  summarize(bv1Frequency = sum(Frequency)) %>%
  select(vb1Value = MaxPMExposure, bv1Frequency)
bv2_df <- bivar_REvsME_df %>%
  group_by(MaxPMExposure) %>%
  summarize(bv2Frequency = sum(Frequency)) %>%
  select(vb2Value = MaxPMExposure, bv2Frequency)
tr1_df <- tri_df %>%
  group_by(MaxPMExposure) %>%
  summarize(triFrequency = sum(Frequency)) %>%
  select(truValue = MaxPMExposure, triFrequency)
un_df <- uni_df %>%
  filter(Feature == "MaxPMExposure") %>%
  rename(uniFrequency = Frequency)
comp3_df <- cbind(un_df,bv1_df,bv2_df,tr1_df) %>%
  select(Feature, Value, uniFrequency, bv1Frequency, bv2Frequency, triFrequency)
comp3_df
```

## Bivariate tables

In this section we compare the three bivariate tables as directly called via the OpenAPI with the marginals of the trivariate table.