---
title: "Validation of Tables"
author: "Haaland"
date: "03/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## Load the tables

In files 2-4, we created and saved the univariate, bivariate and trivariate tables for the variables:

* `EDVisits` = Total ED/Inpatient Visits, 
* `MaxPMExposure` = Maximum Daily PM 2.5 Exposure (qcut), 
* `RoadwayExposure` = Major Roadway Highway Exposure.

The three pairwise combinations of binary tables are labeled as follows:

* `bivar_EDvsME_df`: `EDVisits` versus `MaxPMExposure`
* `bivar_EDvsRE_df`: `EDVisits` versus `RoadwayExposure`
* `bivar_REvsME_df`: `RoadwayExposure` versus `MaxPMExposure`

We start by loading the data.

```{r}
load("RData/uni_df.RData")
load("RData/bivar_EDvsME_df.RData")
load("RData/bivar_EDvsRE_df.RData")
load("RData/bivar_REvsME_df.RData")
load("RData/tri_df.RData")
```



## Brief examination of bivariate tables

The first result that surprises me is that the bivariate table returns fequency values for cells which have counts below 10. I thought that 10 was the minimum number required for reporting.

```{r}
bivar_EDvsME_df %>%
  arrange(Frequency)
```

Compare the total frequencies reported in the univariate table versus the different bivariate tables

```{r}
uni_df %>% 
  group_by(Feature) %>%
  summarize(total_counts = sum(Frequency))

```

```{r}
sum(bivar_EDvsME_df$Frequency)
```
```{r}
uni_df %>% 
  filter(str_detect(Feature, 'EDVisits')) %>%
  filter(as.numeric(Value) >= 10) %>%
  summarize(total_counts = sum(Frequency))
```

We can see that the 57 subjects with 10 or more visits are correctly left out of the bivariate tables based on the total counts.

## Compare bivariate marginals to univariate

To verify the correctness of the data, we can compare the marginal distributions from the bivariate tables to the univariate distributions.

### EDVisits

First we consider the univariate and marginal distributions for EDVisits. We calculate the marginal distributions from the bivariate and trivariate tables and compare to the univariate values. The results are be the same.

```{r}
bv1_df <- bivar_EDvsME_df %>%
  group_by(EDVisits) %>%
  summarize(bv1Frequency = sum(Frequency)) %>%
  select(vb1Value = EDVisits, bv1Frequency)
bv2_df <- bivar_EDvsRE_df %>%
  group_by(EDVisits) %>%
  summarize(bv2Frequency = sum(Frequency)) %>%
  select(vb2Value = EDVisits, bv2Frequency)
tr1_df <- tri_df %>%
  group_by(EDVisits) %>%
  summarize(triFrequency = sum(Frequency)) %>%
  select(truValue = EDVisits, triFrequency)
un_df <- uni_df %>%
  filter(Feature == "EDVisits",
         as.numeric(Value) < 10) %>%
  rename(uniFrequency = Frequency)
comp1_df <- cbind(un_df,bv1_df,bv2_df,tr1_df) %>%
  select(Feature, Value, uniFrequency, bv1Frequency, bv2Frequency, triFrequency)
comp1_df
```

We can see by inspection, that the results look correct. We test that formally below:

```{r}
comp1_df %>%
  transmute(Feature = Feature, Value = Value, agree = ((uniFrequency == bv1Frequency) & 
                                                         (uniFrequency == bv2Frequency) &
                                                        (uniFrequency == triFrequency)))
```

### RoadwayExposure

First we consider the univariate and marginal distributions for RoadwayExposure. We calculate the marginal distributions from the bivariate and trivariate tables and compare to the univariate values. In this comparison, we see differences among the data extraction approaches.

```{r}
bv1_df <- bivar_EDvsRE_df %>%
  group_by(RoadwayExposure) %>%
  summarize(bv1Frequency = sum(Frequency)) %>%
  select(vb1Value = RoadwayExposure, bv1Frequency)
bv2_df <- bivar_REvsME_df %>%
  group_by(RoadwayExposure) %>%
  summarize(bv2Frequency = sum(Frequency)) %>%
  select(vb2Value = RoadwayExposure, bv2Frequency)
tr1_df <- tri_df %>%
  group_by(RoadwayExposure) %>%
  summarize(triFrequency = sum(Frequency)) %>%
  select(truValue = RoadwayExposure, triFrequency)
un_df <- uni_df %>%
  filter(Feature == "RoadwayExposure") %>%
  rename(uniFrequency = Frequency)
comp2_df <- cbind(un_df,bv1_df,bv2_df,tr1_df) %>%
  select(Feature, Value, uniFrequency, bv1Frequency, bv2Frequency, triFrequency)
comp2_df
```

### MaxPMExposure

First we consider the univariate and marginal distributions for MaxPMExposure. We calculate the marginal distributions from the bivariate and trivariate tables and compare to the univariate values. In this comparison, we see differences among the data extraction approaches.

```{r}
bv1_df <- bivar_EDvsME_df %>%
  group_by(MaxPMExposure) %>%
  summarize(bv1Frequency = sum(Frequency)) %>%
  select(vb1Value = MaxPMExposure, bv1Frequency)
bv2_df <- bivar_REvsME_df %>%
  group_by(MaxPMExposure) %>%
  summarize(bv2Frequency = sum(Frequency)) %>%
  select(vb2Value = MaxPMExposure, bv2Frequency)
tr1_df <- tri_df %>%
  group_by(MaxPMExposure) %>%
  summarize(triFrequency = sum(Frequency)) %>%
  select(truValue = MaxPMExposure, triFrequency)
un_df <- uni_df %>%
  filter(Feature == "MaxPMExposure") %>%
  rename(uniFrequency = Frequency)
comp3_df <- cbind(un_df,bv1_df,bv2_df,tr1_df) %>%
  select(Feature, Value, uniFrequency, bv1Frequency, bv2Frequency, triFrequency)
comp3_df
```